---
title: "PA1_template.Rmd"
output: html_document
---

Load and preprocessing the data

``` {r load&preprocess}
setwd("C:/Users/Jack Kerguenel/Documents/Data science/Rep Research/Course Project 1")
## Read the data
rawData<-read.csv("activity.csv")
## Change the Date column from character type to date type
rawData$date<-as.Date(rawData$date, format="%Y-%m-%d")
```

Total number of steps taken per day

```{r TotalNumberOfStepsPerDay}
## Total number of steps taken per day
totalStepsPerDay<-aggregate(rawData$steps,by=list(Group.date=rawData$date),sum)
names(totalStepsPerDay)<-c("Group.date","Total.steps")
```

Histogram of total numbers of steps per day

```{r histogram}
## Parameters og graph
par(las=2,cex=0.5)
## Histogram
barplot(totalStepsPerDay$Total.steps,names.arg=totalStepsPerDay$Group.date)
```

Calculation of the mean and median of the total number of steps taken each day

```{r mean&median}
meanStepsPerDay<-mean(totalStepsPerDay$Total.steps,na.rm=TRUE)
medianStepsPerDay<-median(totalStepsPerDay$Total.steps,na.rm=TRUE)
```

The mean of the total number of steps taken each day is `r meanStepsPerDay` and the median is `r medianStepsPerDay`

Time series plot for mean steps per interval throughout all days

```{r PlotMeansPerInterval}
meanStepsPerInterval<-aggregate(rawData$steps,by=list(Group.interval=rawData$interval),function(x) mean(x,na.rm=TRUE))
names(meanStepsPerInterval)<-c("Group.interval","Mean.steps")
plot(meanStepsPerInterval$Group.interval,meanStepsPerInterval$Mean.steps,type = "l",xlab="Interval",ylab="Mean steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

``` {r WhichMaxIn5MinInterval}
maxIndex<-which.max(meanStepsPerInterval$Mean.steps)
maxInterval<-meanStepsPerInterval$Group.interval[maxIndex]
```
The 5-minute interval that on average contains the maximum number of steps is `r maxInterval`

Number of NAs in the raw data

``` {r NumberOfNAs}
logicVector<-is.na(rawData$steps)
NANumber<-sum(logicVector)
NANumberPercent<-mean(logicVector)*100
```
The number of NAs in the steps column of the raw data is `r NANumber`

Imputing steps with NAs in raw data with the mean of the steps throughout all days in the corresponding interval
```{r Imputting}
rawDataExtended<-data.frame(rawData, is.na=logicVector)
for (i in 1:17568) {
        if (rawDataExtended$is.na[i]==TRUE) {
                interval<-rawDataExtended$interval[i]
                index<-which(meanStepsPerInterval$Group.interval==interval)
                rawDataExtended$steps[i]<-meanStepsPerInterval$Mean.steps[index]
        }
}
```

Total number of steps taken per day with the imputed dataset

```{r TotalNumberOfStepsPerDayWithImputedDataset}
## Total number of steps taken per day
totalStepsPerDayImputed<-aggregate(rawDataExtended$steps,by=list(Group.date=rawDataExtended$date),sum)
names(totalStepsPerDayImputed)<-c("Group.date","Total.steps")
```

Histogram of total numbers of steps per day with the imputed dataset

```{r HistogramImputedDataset}
## Parameters og graph
par(las=2,cex=0.5)
## Histogram
barplot(totalStepsPerDayImputed$Total.steps,names.arg=totalStepsPerDayImputed$Group.date)
```

Calculation of the mean and median of the total number of steps taken each day with the imputed dataset

```{r mean&medianImputedDataset}
meanStepsPerDayImputed<-mean(totalStepsPerDayImputed$Total.steps,na.rm=TRUE)
medianStepsPerDayImputed<-median(totalStepsPerDayImputed$Total.steps,na.rm=TRUE)
```

The mean of the total number of steps taken each day with the imputed dataset is `r meanStepsPerDayImputed` and the median is `r medianStepsPerDayImputed`; as it can be seen the mean does not change from the raw dataset with NAs to the imputed dataset, and the median changes just a bit, mainly because the imputed values are calculated from the 5-minutes intervals means of the rest of the data, which overall makes it more or less the same for summary statistics purposes. 

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day

```{r addingTheFactorWeekendOrWeekday}
## Adding the name of the day
rawDataExtended1<-data.frame(rawDataExtended,weekday=weekdays(rawDataExtended$date),weekdayOrWeekendCharacter=c("weekday"))
## Find out if a day is weekday or weekend and add that as a column of factors
for (i in 1:nrow(rawDataExtended1)) {
        if (rawDataExtended1$weekday[i]=="sábado" | rawDataExtended1$weekday[i]=="domingo") {
        rawDataExtended1$weekdayOrWeekendCharacter[i]<-c("weekend")
        }
}
rawDataExtended1<-data.frame(rawDataExtended1,weekdayOrWeekendFactor=as.factor(rawDataExtended1$weekdayOrWeekendCharacter))
## Eliminate columns that are not useful anymore
rawDataExtended1$is.na=NULL
rawDataExtended1$weekdayOrWeekendCharacter=NULL
```

Panel plot time series plot for mean steps per interval for weekends and for weekdays 

```{r PlotMeansPerIntervalWeekends/Weekdays}
## Subset of weekdays and weekends
weekdaysSubset<-subset(rawDataExtended1,rawDataExtended1$weekdayOrWeekendFactor=="weekday")
weekendSubset<-subset(rawDataExtended1,rawDataExtended1$weekdayOrWeekendFactor=="weekend")
## Average steps for intervals on weekdays
meanStepsPerIntervalWeekdays<-aggregate(weekdaysSubset$steps,by=list(Group.interval=weekdaysSubset$interval),function(x) mean(x,na.rm=TRUE))
names(meanStepsPerIntervalWeekdays)<-c("Interval","Mean steps")
meanStepsPerIntervalWeekdays<-data.frame(meanStepsPerIntervalWeekdays,Weekday.or.weekend=as.factor("weekday"))
## Average steps for intervals on weekends
meanStepsPerIntervalWeekends<-aggregate(weekendSubset$steps,by=list(Group.interval=weekendSubset$interval),function(x) mean(x,na.rm=TRUE))
names(meanStepsPerIntervalWeekends)<-c("Interval","Mean steps")
meanStepsPerIntervalWeekends<-data.frame(meanStepsPerIntervalWeekends,Weekday.or.weekend=as.factor("weekend"))
## Append Mean Steps Per Intervals data frames
mergedDataFrames<-rbind(meanStepsPerIntervalWeekdays,meanStepsPerIntervalWeekends)
names(mergedDataFrames)<-c("Interval","Mean.steps","Weekday.or.weekend")
## Panel plots
library(ggplot2)
graph<-ggplot(mergedDataFrames,aes(x=Interval,y=Mean.steps))
graph<-graph + geom_line() + facet_grid(Weekday.or.weekend ~ .)
graph
```
